Local to global N input modelling

### 2025-06-04 _manuscript.qmd
Move manuscript files to root for deployment

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

@@

```{r global-variables}
#| include: false
#| eval: true

@@

[This is a note create with the `aside` class.]{.aside}

@@

# Methods

@@

farms_geometry <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread() %>% 
  rename(farm_ID = farm_id)
```

@@

```{r fig-farm-mean-temps}
#| fig-cap: |
#|   Potential options for showing farm site mean temperatures. Top left: Mean temperature by latitude across countries. Top right: Frequency distribution of mean temperatures by hemisphere. Bottom: global distribution of mean temperatures.
#| fig-height: 6
#| fig-width: 6

@@

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

@@

## Model and approach

@@

```{r fig-functional-response-to-temperature}
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

@@

* Salmon cohorts were always grown for 548 days
* Salmon cohorts were put out in annually Spring, on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` when farm latitude < 0 and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` when farm latitude > 0.
* Cohorts were put out every year, such that there was an annual overlap of 183 days where one cohort was nearing harvest and another was being initialised.

@@

Table @tbl-params shows the parameter values used to model the growth of farmed Atlantic salmon.

@@

### Testing parameter sensitivity

@@

## Modelled experimental feeds

* Put details of the feeds here - they came from [@cottrell_origins_nodate]
* Key part is the digestibility of the ingredients and feeds overall
* @cottrell_global_2020

@@

New method is to take the average predicted individual harvest size for each farm, then working backwards with the standard mortality rate to generate a population for each farm that will produce the correct tonnage for each farm. @fig-harvest-size shows the  average individual harvest size at each farm as predicted by the fish-growth model, which was used to create the population inputs for the farm-level growth curves.

```{r fig-harvest-size}
#| fig-cap: Frequency distribution of harvest sizes across global farms, as predicted by the fish growth model

file.path(output_farm_data_path, "farm_harvest_size.qs") %>% 
  qread() %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  theme_classic() +
  theme(
    legend.position = "none", 
    aspect.ratio = 0.75
  )
```

This means that the population estimates aren't actual fish numbers, they're just a method that allows each farm location to approximate of the correct tonnage produced while to allowing comparisons between the different feeds.

## Biomass produced

```{r biomass produced}
#| fig-cap: Modelled vs observed salmon biomass produced

coho_biom <- qs::qread(file.path(data_analysis_path, "biomass_produced_comparison.qs"))

@@

## Species vulnerability layers

@@

@@

fnms <- farmrun_reference_files <- file.path(output_cohort_growth_data_path) %>% 
  list.files(full.names = T)

# qs::qread(fnms[1]) %>% names()
weight <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["weight_stat"]]})
biomass <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["biomass_stat"]]})

# The following reads in all the total excretion data, pairs it with the farm coordinates, and converts the mean and sd values into values/biomass
excr_total <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["total_excr_stat"]]}) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  mutate(sd.x = sd.x/mean.x,
         sd.y = sd.x/mean.y,
         mean.x = set_units(mean.x, "g d-1"),
         mean.y = set_units(mean.y, "gfish"),
         mean = mean.x/mean.y,
         sd_pc = sqrt(sd.x^2+sd.y^2),
         sd = sd_pc * mean,
         prod_day = case_when(hemisphere == "S" ~ t-(274+365)+1, T ~ t-(121+365)+1)) %>% 
  select(-c(t_start, t_end, mean.x, mean.y, sd.x, sd.y))

excr_P <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["P_excr_stat"]]}) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  mutate(sd.x = sd.x/mean.x,
         sd.y = sd.x/mean.y,
         mean.x = set_units(mean.x, "g d-1"),
         mean.y = set_units(mean.y, "gfish"),
         mean = mean.x/mean.y,
         sd_pc = sqrt(sd.x^2+sd.y^2),
         sd = sd_pc * mean) %>% 
  select(-c(t_start, t_end, mean.x, mean.y, sd.x, sd.y))

excr_L <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["L_excr_stat"]]}) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  mutate(sd.x = sd.x/mean.x,
         sd.y = sd.x/mean.y,
         mean.x = set_units(mean.x, "g d-1"),
         mean.y = set_units(mean.y, "gfish"),
         mean = mean.x/mean.y,
         sd_pc = sqrt(sd.x^2+sd.y^2),
         sd = sd_pc * mean) %>% 
  select(-c(t_start, t_end, mean.x, mean.y, sd.x, sd.y))

excr_C <- purrr::map_dfr(fnms, function(fnm) {qs::qread(fnm)[["C_excr_stat"]]}) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  mutate(sd.x = sd.x/mean.x,
         sd.y = sd.x/mean.y,
         mean.x = set_units(mean.x, "g d-1"),
         mean.y = set_units(mean.y, "gfish"),
         mean = mean.x/mean.y,
         sd_pc = sqrt(sd.x^2+sd.y^2),
         sd = sd_pc * mean) %>% 
  select(-c(t_start, t_end, mean.x, mean.y, sd.x, sd.y))
```

### Total amount

@@

df <- excr_total %>% 
  select(-c(sd, sd_pc)) %>% 
  pivot_wider(names_from = "feed", values_from = "mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kgfish-1 d-1"),
         diff_perc = drop_units(diff/reference))

@@

df <- excr_total %>% 
  select(-c(sd, sd_pc)) %>% 
  pivot_wider(names_from = "feed", values_from = "mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kgfish-1 d-1"),
         diff = set_units(diff, "kg tfish-1 d-1"))

@@

## Uneaten feed (amount & composition)

>>>>

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

@@

```{r global-variables}
#| include: false

@@

[This intro is copied verbatim from work done during one of the early workshops.]{.aside}

@@

# Methods
## Model and approach

[Unfinished. Need to work out how much should be transposed from @baldan_r_2018. One thing that definitely needs to be stated is the individual ingredient (rather than whole food) tracking, the use of cohorts, and the new method for calculating population.]{.aside}

@@

farms_geometry <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread() %>% 
  rename(farm_ID = farm_id)

sorted_countries <- farm_coords %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  group_by(country) %>% 
  reframe(av_lat = mean(lat)) %>% 
  arrange(-av_lat) %>% 
  pull(country) %>% 
  unique()
```

@@

```{r fig-farm-mean-temps}
#| fig-cap: |
#|   Potential options for showing farm site mean temperatures. Top left: Mean temperature by latitude across countries. Top right: Frequency distribution of mean temperatures by hemisphere. Bottom: global distribution of mean temperatures.
#| fig-height: 9
#| fig-width: 9.5

@@

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

@@

@@

To detail here:

* Functional response to temperature (species-specific)
* Overfeeding parameters and how they come into play
* Deminstration of how feeding response and actual food given vary with temperature?

```{r fig-functional-response-to-temperature}
#| eval: false
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

@@

* Salmon cohorts were always grown for 548 days
* Salmon cohorts were initialised Spring. This was on day `r unique(farm_coords$t_start[farm_coords$lat < 0])` for farms in the southern hemisphere and day `r unique(farm_coords$t_start[farm_coords$lat > 0])` for farms in the northern hemisphere.
* Cohorts were initialised anually, such that there was an overlap of 183 days where one cohort was nearing harvest and another was being initialised.

@@

tbl-params shows the parameter values used to model the growth of farmed Atlantic salmon.

@@

### Parameter sensitivity

@@

## Modelled experimental feeds {#sec-feeds}

```{r feeds}
feed_pal <- c("past" = brewer.pal(9, "Set1")[1], "reference" = brewer.pal(9, "Set1")[2], "future" = brewer.pal(9, "Set1")[3])
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>% qs::qread()

df <- feed_params[["reference"]]
contribP_1 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_1 <- sum(contribP_1)
digestP_1 <- sum(df[["Proteins"]]$digest * contribP_1)/tP_1

contribL_1 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_1 <- sum(contribL_1)
digestL <- sum(df[["Lipids"]]$digest * contribL_1)/tL_1

contribC_1 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_1 <- sum(contribC_1)
digestC_1 <- sum(df[["Carbohydrates"]]$digest * contribC_1)/tC_1

df <- feed_params[["past"]]
contribP_2 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_2 <- sum(contribP_2)
digestP_2 <- sum(df[["Proteins"]]$digest * contribP_2)/tP_2

contribL_2 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_2 <- sum(contribL_2)
digestL <- sum(df[["Lipids"]]$digest * contribL_2)/tL_2

contribC_2 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_2 <- sum(contribC_2)
digestC_2 <- sum(df[["Carbohydrates"]]$digest * contribC_2)/tC_2

df <- feed_params[["future"]]
contribP_3 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_3 <- sum(contribP_3)
digestP_3 <- sum(df[["Proteins"]]$digest * contribP_3)/tP_3

contribL_3 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_3 <- sum(contribL_3)
digestL <- sum(df[["Lipids"]]$digest * contribL_3)/tL_3

contribC_3 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_3 <- sum(contribC_3)
digestC_3 <- sum(df[["Carbohydrates"]]$digest * contribC_3)/tC_3
```

[Possibly replace this with a better summary of the ingredients that make up each feed, and report these stats in Results under Uneaten feed]{.aside}

* Feed came from [@cottrell_origins_nodate; @cottrell_global_2020]
* The reference feed is `r round(100*tP_1,1)`% protein, `r round(100*tL_1,1)`% lipid, `r round(100*tC_1,1)`% carbohydrate. 
* The past feed is `r round(100*tP_2,1)`% protein, `r round(100*tL_2,1)`% lipid, `r round(100*tC_2,1)`% carbohydrate. 
* The future feed is `r round(100*tP_3,1)`% protein, `r round(100*tL_3,1)`% lipid, `r round(100*tC_3,1)`% carbohydrate. 
* Within the reference feed, protein is overall `r round(100*digestP_1,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_1,1)`% digestible.
* Within the reference feed, protein is overall `r round(100*digestP_2,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_2,1)`% digestible.
* Within the reference feed, protein is overall `r round(100*digestP_3,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_3,1)`% digestible.

@@

```{r pop-biom-data}
coho_biom <- file.path(data_analysis_path, "biomass_produced_comparison.qs") %>% 
  qs::qread()
harv_size <- file.path(output_farm_data_path, "farm_harvest_size.qs") %>% 
  qread()

coho_biom_summ <- coho_biom %>% 
  mutate(perc_diff = 100*(mean - tonnes_per_farm)/tonnes_per_farm)
```

Each farm's final population was calculated by dividing the production tonnage required by the average modelled individual harvest size for each farm (when feed reference feed). A population timeseries was the generated using the standard mortality rate (`r unname(params['mortmyt'])` d$^{-1}$). Using this method, a farm's 'population' is not a good representation of the actual number of individuals being farmed, but it ensures each farm location is producing the correct tonnage while allowing comparisons between the different feeds.

@fig-harvest-size shows the distribution of average individual harvest sizes across farms used to create the population timeseries as predicted by the fish-growth model.

```{r fig-harvest-size, class.chunk = "aside"}
#| fig-cap: Frequency distribution of harvest sizes across global farms, as predicted by the fish growth model

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  theme_classic() +
  theme(
    legend.position = "none", 
    aspect.ratio = 0.75
  )
```

By this method, farm biomass as predicted by the model slightly overestimated farm tonnage. Predicted farm biomass was `r round(min(coho_biom_summ$perc_diff),1)`-`r round(max(coho_biom_summ$perc_diff),1)` higher than tonnage produced (@fig-biomass-produced). This variation was not correlated with mean farm temperature or total tonnage produced (@fig-biomass-variation).

```{r fig-biomass-variation, class.chunk = "aside"}
plot(coho_biom_summ$temp_c, coho_biom_summ$perc_diff)
plot(coho_biom_summ$tonnes_per_farm, coho_biom_summ$perc_diff)
```

```{r fig-biomass-produced, class.chunk = "aside"}
#| fig-cap: Modelled vs observed salmon biomass produced

@@

## Impacts of nutrient input

*To detail here:*

* How to convert feed and uneaten faeces into nutrients - just nitrogen? Carbon too? If just nitrogen, then can probably use the standard N = 6.25 x protein conversion to compare the feeds (lipids and carbohydrates have negligible protein). If doing carbon too, will need to find conversion rates for all of them, possibly splitting carbs/lipids into complex/simple, saturated/unsaturated, etc.
* Source of species vulnerability layers

@@

```{r weight and biomass data}
farmrun_files <- file.path(output_cohort_growth_data_path) %>% 
  list.files(full.names = T)

qs::qread(farmrun_files[1]) %>% names()
weight <- purrr::map(farmrun_files, function(f) {qs::qread(f)[["weight_stat"]]}) %>% bind_rows()
biomass <- purrr::map(farmrun_files, function(f) {qs::qread(f)[["biomass_stat"]]}) %>% bind_rows()
```

## Uneaten feed

```{r uneaten feed data}
#| code-summary: Load and prep uneaten feed data

uneaten_total <- farmrun_files %>% 
  purrr::map(function(f) {qs::qread(f)[["total_uneat_stat"]]}) %>% 
  bind_rows() %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(uneat_mean = mean.x, biomass_mean = mean.y,
         uneat_sd = sd.x, biomass_sd = sd.y) %>% 
  mutate(uneat_sd = uneat_sd/uneat_mean,
         biomass_sd = biomass_sd/biomass_mean,
         uneat_biom_mean = uneat_mean/biomass_mean,
         uneat_biom_sd = sqrt(uneat_sd^2+biomass_sd^2)*uneat_biom_mean)

uneaten_total_stats <- uneaten_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(uneat_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g gfish-1") %>% set_units("g kgfish-1"),
         max = max %>% set_units("g gfish-1") %>% set_units("g kgfish-1"),
         mean = mean %>% set_units("g gfish-1") %>% set_units("g kgfish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))
```

### Amount

* Total reference uneaten feed across a cohort (548 days) ranged from `r uneaten_total_stats$min[uneaten_total_stats$feed == "reference"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "reference"]` g kg$^{-1}$, with a mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "reference"]`.
* There was very little difference between the past and future feeds and the reference feed. Uneaten past feed ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "past"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "past"]` g kg$^{-1}$ (mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "past"]`) while uneaten future feed ranged from 
`r uneaten_total_stats$min[uneaten_total_stats$feed == "future"]` to `r uneaten_total_stats$min[uneaten_total_stats$feed == "future"]` g kg$^{-1}$ (mean of `r uneaten_total_stats$mean[uneaten_total_stats$feed == "future"]`).
* @fig-total-uneaten-perday shows the range of total uneaten feed per day across countries.

```{r fig-total-uneaten-perday}
#| fig-width: 9.5

uneaten_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(uneat_biom_mean = uneat_biom_mean %>% set_units("g gfish-1 d-1") %>% set_units("g kgfish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = uneat_biom_mean, fill = feed)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Uneaten feed", x = "Country")
```

```{r uneaten composition}
#| eval: false

uneaten_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_uneat_stat"]]}) %>% bind_rows()
uneaten_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_uneat_stat"]]}) %>% bind_rows()
uneaten_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_uneat_stat"]]}) %>% bind_rows()

uneaten_all <- uneaten_total %>% 
  select(-contains("biom")) %>% 
  merge(uneaten_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(uneaten_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd")))
```

* The composition of the uneaten feed is (obviously) the same as the total feed composition (@sec-feeds).

@@

# The following reads in all the total excretion data and converts the mean and sd values into values/biomass
excreted_total <- farmrun_files %>% 
  purrr::map(function(f) {qs::qread(f)[["total_excr_stat"]]}) %>% 
  bind_rows() %>% 
  merge(biomass, by = c("farm_ID", "feed", "t")) %>% 
  rename(excr_mean = mean.x, biomass_mean = mean.y,
         excr_sd = sd.x, biomass_sd = sd.y) %>% 
  mutate(excr_sd = excr_sd/excr_mean,
         biomass_sd = biomass_sd/biomass_mean,
         excr_biom_mean = excr_mean/biomass_mean,
         excr_biom_sd = sqrt(excr_sd^2+biomass_sd^2)*excr_biom_mean)

excreted_total_stats <- excreted_total %>% 
  group_by(feed, farm_ID) %>% 
  reframe(total = sum(excr_biom_mean)) %>% 
  group_by(feed) %>% 
  reframe(min = minna(total),
          max = maxna(total),
          mean = meanna(total)) %>% 
  mutate(min = min %>% set_units("g gfish-1") %>% set_units("g kgfish-1"),
         max = max %>% set_units("g gfish-1") %>% set_units("g kgfish-1"),
         mean = mean %>% set_units("g gfish-1") %>% set_units("g kgfish-1")) %>% 
  mutate(min = round(min,2),
         max = round(max,2),
         mean = round(mean,2))
```

* Total reference excreted faeces across a cohort (548 days) ranged from `r excreted_total_stats$min[excreted_total_stats$feed == "reference"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "reference"]` g kg$^{-1}$, with a mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "reference"]`.
* There was very little difference between the past and future feeds and the reference feed. Uneaten past feed ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "past"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "past"]` g kg$^{-1}$ (mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "past"]` g kg$^{-1}$) while uneaten future feed ranged from 
`r excreted_total_stats$min[excreted_total_stats$feed == "future"]` to `r excreted_total_stats$max[excreted_total_stats$feed == "future"]` g kg$^{-1}$ (mean of `r excreted_total_stats$mean[excreted_total_stats$feed == "future"]` g kg$^{-1}$).
* @fig-total-excreted-perday shows the range of total excreted faeces per day across countries.

```{r fig-total-excreted-perday}
#| fig-width: 9

excreted_total %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  mutate(excr_biom_mean = excr_biom_mean %>% set_units("g gfish-1 d-1") %>% set_units("g kgfish-1 d-1"),
         country = factor(country, levels = sorted_countries)) %>% 
  ggplot(aes(x = country, y = excr_biom_mean, fill = feed)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Excreted faeces", x = "Country")
```

### Amount

@@

df <- excreted_total %>% 
  select(-contains(c("sd", "biom"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kgfish-1 d-1"),
         diff_perc = drop_units(diff/reference)) %>% 
    merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t - t_start + 1)

@@

df <- excreted_total %>% 
  select(-contains(c("sd", "excr_mean", "biomass"))) %>% 
  pivot_wider(names_from = "feed", values_from = "excr_biom_mean") %>% 
  mutate(past = past-reference,
         future = future-reference) %>% 
  pivot_longer(names_to = "feed", values_to = "diff", cols = c(past, future), names_transform = list(feed = as.factor)) %>% 
  mutate(reference = set_units(reference, "g kgfish-1 d-1"),
         diff = set_units(diff, "kg tfish-1 d-1")) %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  mutate(prod_day = t-t_start+1)

@@

### Composition

```{r excreted composition}
#| eval: true

excreted_P <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["P_excr_stat"]]}) %>% bind_rows()
excreted_L <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["L_excr_stat"]]}) %>% bind_rows()
excreted_C <- farmrun_files %>% purrr::map(function(f) {qs::qread(f)[["C_excr_stat"]]}) %>% bind_rows()

excreted_all <- excreted_total %>% 
  select(-contains("biom")) %>% 
  merge(excreted_P, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_L, by = c("farm_ID", "feed", "t")) %>% 
  merge(excreted_C, by = c("farm_ID", "feed", "t")) %>% 
  select(-c(contains("sd"))) %>% 
  rename(protein = mean.x, lipids = mean.y, carbohydrates = mean) %>% 
  mutate(protein = protein/excr_mean %>% set_units("d-1"),
         lipids = lipids/excr_mean %>% set_units("d-1"),
         carbohydrates = carbohydrates/excr_mean %>% set_units("d-1")) %>% 
  select(-excr_mean)
```

Unlike the uneaten feed, the composition of excreted faeces differs from that of the incoming feed.

* Composition does not change through time or with increasing biomass (because digestibility does not change with fish size or temeprature)

```{r composition-throughtime}
#| eval: false

excreted_all %>% 
  pivot_longer(names_to = "macro", values_to = "perc", cols = c("protein", "lipids", "carbohydrates")) %>% 
  mutate(macro = factor(macro, levels = c("protein", "lipids", "carbohydrates"))) %>% 
  group_by(farm_ID, feed, macro) %>% 
  reframe(sd = sd(perc) %>% set_units("d-1"),
          mean = mean(perc)) %>% 
  ggplot(aes(x = farm_ID, colour = feed, fill = feed, y = 100*mean, ymin = 100*(mean-sd), ymax = 100*(mean+sd))) +
  geom_point(size = 2) +
  geom_errorbar(width = 0.5) +
  scale_color_manual(values = feed_pal) +
  scale_fill_manual(values = feed_pal) +
  facet_grid(rows = vars(macro)) +
  theme_classic() +
  labs(x = "Macronutrient", y = "Mean composition of faeces")
```
